---
title: "topONE"
author: "Claire"
date: "2024-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# tinytex::install_tinytex()
```

## Setup

```{r}
library(dplyr)
library(car)
library(FSA)
```

```{r}
df <- read.csv("Documents/PGC Bioinfo Training/Viral Recombination/Topo ARG/country_homologies.csv")
attach(df); View(df)
```

We take manual setting of data frames per country.

```{r}
ph_df <- df %>%
  filter(country_code == 'PH', homology == 1) %>%
  select(gene_type, b_time, d_time)

cn_df <- df %>%
  filter(country_code == 'CN', homology == 1) %>%
  select(gene_type, b_time, d_time)

sg_df <- df %>%
  filter(country_code == 'SG', homology == 1) %>%
  select(gene_type, b_time, d_time)

sk_df <- df %>%
  filter(country_code == 'SK', homology == 1) %>%
  select(gene_type, b_time, d_time)

us_df <- df %>%
  filter(country_code == 'US', homology == 1) %>%
  select(gene_type, b_time, d_time)
```

Try input them into one list for brevity

```{r}
country_dfs <- list()
country_dfs[['PH']] <- ph_df
country_dfs[['CN']] <- cn_df
country_dfs[['SG']] <- sg_df
country_dfs[['SK']] <- sk_df
country_dfs[['US']] <- us_df
```

# Philippines

## Births

Test assumptions on the anova model

```{r}
model_ph_births <- aov(b_time ~ gene_type, ph_df)
# homogeneity of variance
leveneTest(model_ph_births) #pval>0.05 --> homoscedastic

# normality of residuals
resid_ph <- residuals(model_ph_births)
shapiro.test(resid_ph) #p-val<0.05 i.e., not normal
```

-   For the test on homogeneity of variance, at \$p\$-val \$=0.3381 \> 0.05\$, we fail to reject the null hypothesis, i.e. the model is homoscedastic.

-   With a Shapiro-Wilk normality test, at p-val \< 0.05, we reject the null hypothesis, i.e., the residuals are not normally distributed.

Since the assumptions are not satisfied, we proceed with a nonparametric Kruskal-Wallis test.

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(b_time ~ gene_type, ph_df)
# p-val = 0.0489 < 0.05 --> there is a significant difference in birth times between gene types

# post-hoc
dunnTest(b_time ~ gene_type, data=ph_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 0.0489 \< 0.05 so we reject the null hypothesis. That is, there is a significant difference in birth times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> significant

-   Mixed vs Recombinant —\> not significant

-   Non-recombinant vs Recombinant —\> not significant

\*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(b_time~gene_type, data=ph_df, ylab="Birth Times", main="Birth Times of SARS-CoV-2 in the PH across different Gene Types", xlab="Gene Type")
```

## Deaths

Test assumptions on the anova model

```{r}
model_ph_deaths <- aov(d_time ~ gene_type, ph_df)
# homogeneity of variance
leveneTest(model_ph_deaths) #pval>0.05 --> homoscedastic

# normality of residuals
resid_ph2 <- residuals(model_ph_deaths)
shapiro.test(resid_ph2) #p-val<0.05 i.e., not normal
```

-   For the test on homogeneity of variance, at \$p\$-val \$=0.5452 \> 0.05\$, we fail to reject the null hypothesis, i.e. the model is homoscedastic.

-   With a Shapiro-Wilk normality test, at p-val \< 0.05, we reject the null hypothesis, i.e., the residuals are not normally distributed.

Since the assumptions are not satisfied, we proceed with a nonparametric Kruskal-Wallis test.

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(d_time ~ gene_type, ph_df)
# p-val = 0.0489 < 0.05 --> there is a significant difference in birth times between gene types

# post-hoc
dunnTest(d_time ~ gene_type, data=ph_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 0.1425 \> 0.05 so we fail to reject the null hypothesis. That is, there is no significant difference in birth times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> not significant

-   Mixed vs Recombinant —\> not significant

-   Non-recombinant vs Recombinant —\> not significant

\*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(d_time~gene_type, data=ph_df, ylab="Death Times", main="Birth Times of SARS-CoV-2 in the PH across different Gene Types", xlab="Gene Type")
```

## 

# China

## Births

Test assumptions on the anova model

```{r}
model_cn_births <- aov(b_time ~ gene_type, cn_df)
# homogeneity of variance
leveneTest(model_cn_births) #pval>0.05 --> heteroscedastic

# normality of residuals
resid_cn <- residuals(model_cn_births)
shapiro.test(resid_cn) #p-val<0.05 i.e., not normal
```

-   For the test on homogeneity of variance, at \$p\$-val \$ \> 0.05\$, we fail to reject the null hypothesis, i.e. the model is homoscedastic.

-   With a Shapiro-Wilk normality test, at p-val \< 0.05, we reject the null hypothesis, i.e., the residuals are not normally distributed.

Since the assumptions are not satisfied, we proceed with a nonparametric Kruskal-Wallis test.

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(b_time ~ gene_type, cn_df)
# p-val = 1 > 0.05 --> there is no significant difference in birth times between gene types

# post-hoc
dunnTest(b_time ~ gene_type, data=cn_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 0.0489 \< 0.05 so we reject the null hypothesis. That is, there is a significant difference in birth times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> significant

    \*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(b_time~gene_type, data=cn_df, ylab="Birth Times", main="Birth Times of SARS-CoV-2 in the CN across different Gene Types", xlab="Gene Type")
```

## Deaths

Test assumptions on the anova model

```{r}
model_cn_deaths <- aov(d_time ~ gene_type, cn_df)
# homogeneity of variance
leveneTest(model_cn_deaths) #pval>0.05 --> homoscedastic

# normality of residuals
resid_ph3<- residuals(model_cn_deaths)
shapiro.test(resid_ph3) #p-val<0.05 i.e., not normal
```

-   For the test on homogeneity of variance, at \$p\$-val \$\> 0.05\$, we fail to reject the null hypothesis, i.e. the model is homoscedastic.

-   With a Shapiro-Wilk normality test, at p-val \< 0.05, we reject the null hypothesis, i.e., the residuals are not normally distributed.

Since the assumptions are not satisfied, we proceed with a nonparametric Kruskal-Wallis test.

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(d_time ~ gene_type, cn_df)
# p-val = 0.0489 < 0.05 --> there is a significant difference in birth times between gene types

# post-hoc
dunnTest(d_time ~ gene_type, data=cn_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 0.1425 \> 0.05 so we fail to reject the null hypothesis. That is, there is no significant difference in deaths times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> not significant

\*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(d_time~gene_type, data=cn_df, ylab="Death Times", main="Birth Times of SARS-CoV-2 in the PH across different Gene Types", xlab="Gene Type")
```

# Singapore

## Births

Test assumptions on the anova model

```{r}
model_sg_births <- aov(b_time ~ gene_type, sg_df)
# homogeneity of variance
leveneTest(model_sg_births) #pval>0.05 --> heteroscedastic

# normality of residuals
resid_sg <- residuals(model_sg_births)
shapiro.test(resid_sg) #p-val<0.05 i.e., not normal
```

-   For the test on homogeneity of variance, at \$p\$-val \$ \> 0.05\$, we fail to reject the null hypothesis, i.e. the model is homoscedastic.

-   With a Shapiro-Wilk normality test, at p-val \< 0.05, we reject the null hypothesis, i.e., the residuals are not normally distributed.

Since the assumptions are not satisfied, we proceed with a nonparametric Kruskal-Wallis test.

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(b_time ~ gene_type, data = sg_df)
# p-val = 1 > 0.05 --> there is no significant difference in birth times between gene types

# post-hoc
dunnTest(b_time ~ gene_type, data=sg_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 1 \> 0.05 so we fail reject the null hypothesis. That is, there is no significant difference in birth times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> significant

    \*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(b_time~gene_type, data=sg_df, ylab="Birth Times", main="Birth Times of SARS-CoV-2 in SG across different Gene Types", xlab="Gene Type")
```

## Deaths

Test assumptions on the anova model

```{r}
model_sg_deaths <- aov(d_time ~ gene_type, sg_df)
# homogeneity of variance
leveneTest(model_sg_deaths) #pval>0.05 --> homoscedastic

# normality of residuals
resid_sg<- residuals(model_sg_deaths)
shapiro.test(resid_sg) #p-val<0.05 i.e., not normal
```

-   For the test on homogeneity of variance, at \$p\$-val \$\> 0.05\$, we fail to reject the null hypothesis, i.e. the model is homoscedastic.

-   With a Shapiro-Wilk normality test, at p-val \< 0.05, we reject the null hypothesis, i.e., the residuals are not normally distributed.

Since the assumptions are not satisfied, we proceed with a nonparametric Kruskal-Wallis test.

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(d_time ~ gene_type, sg_df)
# p-val = 0.0489 < 0.05 --> there is a significant difference in birth times between gene types

# post-hoc
dunnTest(d_time ~ gene_type, data=sg_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 0.1425 \> 0.05 so we fail to reject the null hypothesis. That is, there is no significant difference in deaths times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> not significant

\*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(d_time~gene_type, data=sg_df, ylab="Death Times", main="Birth Times of SARS-CoV-2 in the SG across different Gene Types", xlab="Gene Type")
```

## 

# South Korea

## Births

(Skipping the assumptions for brevity of file but all are failed)

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(b_time ~ gene_type, data = sk_df)
# p-val = 1 > 0.05 --> there is no significant difference in birth times between gene types

# post-hoc
dunnTest(b_time ~ gene_type, data=sk_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 0.0461 \< 0.05 so we reject the null hypothesis. That is, there is a significant difference in birth times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> not significant

-   Mixed vs Recombinant —\> (not) significant

-   Non-recombinant vs Recombinant —\> (not) significant

\*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(b_time~gene_type, data=sk_df, ylab="Death Times", main="Birth Times of SARS-CoV-2 in SK across different Gene Types", xlab="Gene Type")
```

## Deaths

(Skipping the assumptions for brevity of file but all are failed)

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(d_time ~ gene_type, data = sk_df)
# p-val = 1 > 0.05 --> there is no significant difference in birth times between gene types

# post-hoc
dunnTest(d_time ~ gene_type, data=sk_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 0.04732 \< 0.05 so we reject the null hypothesis. That is, there is a significant difference in birth times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> not significant

-   Mixed vs Recombinant —\> (not) significant

-   Non-recombinant vs Recombinant —\> (not) significant

\*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(d_time~gene_type, data=sk_df, ylab="Death Times", main="Death Times of SARS-CoV-2 in SK across different Gene Types", xlab="Gene Type")
```

## 

# United States

## Births

(Skipping the assumptions for brevity of file but all are failed)

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(b_time ~ gene_type, data = us_df)
# p-val = 1 > 0.05 --> there is no significant difference in birth times between gene types

# post-hoc
dunnTest(b_time ~ gene_type, data=us_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 1\>0.05 so we fail to reject the null hypothesis. That is, there is no significant difference in birth times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> not significant

\*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(b_time~gene_type, data=us_df, ylab="Death Times", main="Birth Times of SARS-CoV-2 in the US across different Gene Types", xlab="Gene Type")
```

## Deaths

(Skipping the assumptions for brevity of file but all are failed)

```{r}
# ... Assumptions failed, nonparametric: Kruskal-Wallis
kruskal.test(d_time ~ gene_type, data = us_df)
# p-val = 1 > 0.05 --> there is no significant difference in birth times between gene types

# post-hoc
dunnTest(d_time ~ gene_type, data=us_df, method="bonferroni") 
#__ all three time pairs are significantly different
```

The Kruskal-Wallis test returns a p-value = 1\> 0.05 so we fail to reject the null hypothesis. That is, there is no significant difference in birth times between the three gene types. Post-hoc analysis shows the following:

-   Mixed vs Nonrecombinant —\> not significant

\*Graphs\*

```{r}
order1 <- c("recombinant", "nonrecombinant", "mixed")
genetype <- factor(gene_type, levels=order1)

boxplot(d_time~gene_type, data=us_df, ylab="Death Times", main="Death Times of SARS-CoV-2 in the US across different Gene Types", xlab="Gene Type")
```

## 

# 
